# Using BigQuery

## Data

We will use the NYC Bicycle Hire data we looked at in lecture. The database contains all Citi Bike trips (NYC's bike-sharing service) since Citi Bike launched in September 2013.

```{r}
library("DBI")
library("bigrquery")
```

## Queries

1. First, connect to the database and see the first 5 entries.

```{r}
db <- dbConnect(
  bigrquery::bigquery(),
  project = "bigquery-public-data",
  dataset = "new_york",
  billing = "my472-week-10-lecture"
)

dbGetQuery(db, "SELECT * FROM citibike_trips LIMIT 5")

```

2. What is the average trip duration based on the age of the riders?

```{r}
dbGetQuery(db, "SELECT birth_year, AVG( TIMESTAMP_DIFF(stoptime, starttime, MINUTE) )
          AS duration_minutes
          FROM citibike_trips
          GROUP BY birth_year
          ORDER BY birth_year")

res <- dbGetQuery(db, "SELECT (2020-birth_year) AS age,
          AVG( TIMESTAMP_DIFF(stoptime, starttime, MINUTE )) AS duration_minutes,
          COUNT(*) AS trip_count
          FROM citibike_trips
          GROUP BY age
          ORDER BY age")
plot(res$age[res$age < 100], res$duration_minutes[res$age < 100], type = 'l')
barplot(res$trip_count[!is.na(res$age) & res$age<100], 
        names.arg=res$age[!is.na(res$age) & res$age<100])
```

3. What is the average distance of a CitiBike trip based on the age of the riders?

```{r}
res <- dbGetQuery(db, "SELECT (EXTRACT(YEAR FROM CURRENT_TIMESTAMP) - birth_year) AS age,
           AVG( (ABS(start_station_latitude - end_station_latitude)
           + ABS(start_station_longitude - end_station_longitude)) * 111) AS average_distance,
           FROM citibike_trips
           GROUP BY age
           ORDER BY age")
plot(res$age[res$age < 100], res$average_distance[res$age < 100], type ='l' )
```

4. What is the average speed of a CitiBike trip?

```{r}
dbGetQuery(db,
           "SELECT AVG( ((ABS(start_station_latitude - end_station_latitude)
           + ABS(start_station_longitude - end_station_longitude)) * 111) / 
           tripduration) AS average_speed
           FROM citibike_trips")

query_exec(
  "SELECT AVG(((ABS(start_station_latitude-end_station_latitude) + 
        ABS(start_station_longitude-end_station_longitude)) * 111) /
        ((stoptime-starttime)/1000000/3600) ) AS avg_speed_km_h
  FROM [bigquery-public-data:new_york.citibike_trips]",
  project = "my472-week-10-lecture") # own project, because no create right for public data
```

5. What is the average speed based on age?

```{r}
dbGetQuery(db,
           "SELECT (EXTRACT(YEAR FROM CURRENT_TIMESTAMP) - birth_year) AS age,
           AVG( (ABS(start_station_latitude - end_station_latitude)
           + ABS(start_station_longitude - end_station_longitude)) * 111 / tripduration )
           AS average_speed
           FROM citibike_trips
           GROUP BY age")

res <- query_exec(
        "SELECT (2020 - birth_year) AS age,
            AVG(((ABS(start_station_latitude-end_station_latitude) + 
            ABS(start_station_longitude-end_station_longitude)) * 111) /
            ((stoptime-starttime)/1000000/3600) ) AS avg_speed_km_h
        FROM [bigquery-public-data:new_york.citibike_trips]
        GROUP BY age
        ORDER BY age",
        project = "my472-week-10-lecture")

plot(res$age[res$age < 100], res$avg_speed_km_h[res$age < 100], type = 'l',
     xlab = 'rider age', ylab = 'average speed in km/h' )

```
