# Creating and querying a SQL database

Practice creating a SQL database and querying it. You should use the data in the file UK-tweets.csv. These data represent a random sample of 10,000 tweets related to the 2014 EP elections in the UK.

1. Create a SQLite database from the data in the file. Test the database by checking the first 5 entries in it.

```{r}
library(DBI)

database <- dbConnect(RSQLite::SQLite(), "facebook-data/UK-tweets-db.sqlite")
# dbRemoveTable(database, "tweets")
tweets <- read.csv("UK-tweets.csv", stringsAsFactors = FALSE)
dbWriteTable(database, "tweets", tweets, overwrite = TRUE)
dbListFields(database, "tweets")
dbGetQuery(database, "SELECT * FROM tweets LIMIT 5")
```

2. How many tweets are classified as engaging? (HINT: Use CASE WHEN)

```{r}
#dbGetQuery(database, )
#dbDisconnect(database)
```

```{sql, connection=database}
SELECT COUNT(*) AS number_engaging
FROM tweets
WHERE communication == "engaging"
```

```{sql, connection=database}
SELECT SUM(CASE WHEN communication='engaging' THEN 1 ELSE 0 END) AS number_engaging 
FROM tweets
```


3. What perecent of tweets are classified as impolite?

```{sql, connection=database}
SELECT 100.0*SUM(CASE WHEN polite = "impolite" THEN 1 ELSE 0 END)/COUNT(*) AS impolite_perc
FROM tweets
```
```{r} 
# Alternaive solution, a bit funny
dbGetQuery(database, 
           "SELECT 100.0*SUM(CASE WHEN polite LIKE '%im%' THEN 1 ELSE NULL END)/COUNT(*) AS Brexit_perc
           FROM tweets")
```



4. What percent of tweets mention Brexit?

```{sql, connection=database}
SELECT 100.0*SUM(CASE WHEN text LIKE '%Brexit%' THEN 1 ELSE NULL END)/COUNT(*) AS brexit_perc
FROM tweets
```


5. Compute the average number of followers for tweets that are impolite and polite.

```{sql, connection = database}
SELECT polite, AVG(followers_count) AS average_followers
FROM tweets
GROUP BY polite

```


```{r}
dbDisconnect(db)
```
