# Practicing working with multiple SQL tables

Practice writing using **JOIN** in SQL queries to answer the questions below. Use the database we created in lecture.

1. How many posts were published each month of the year?

```{r}
library(DBI)
db <- dbConnect(RSQLite::SQLite(), "facebook-data/facebook-db.sqlite")

dbListTables(db)
dbListFields(db, "posts")
dbGetQuery(db,
           "SELECT * FROM posts") # to take a look at posts table

dbGetQuery(db, 
           "SELECT SUBSTR(date, 1, 7) AS month, COUNT(*) AS number_of_posts
           FROM posts
           GROUP BY month")
```


2. What is the average number of likes for members of the House and the Senate? Which group is more popular on average?

```{r}
dbGetQuery(db, 
           "SELECT congress.type, AVG(likes_count) AS avg_likes
           FROM congress JOIN posts
             ON congress.screen_name = posts.screen_name
           GROUP BY congress.type")
```

3. What is the average number of likes that posts receive based on the time of the day in which they were published?

```{r}
res <- dbGetQuery(db,
                  "SELECT SUBSTR(datetime, 12, 2) AS hour_of_day, AVG(likes_count) AS avg_likes
                  FROM posts
                  GROUP BY hour_of_day")

plot(res$hour_of_day, res$avg_likes, type = "l")

```

4. Do women receive more comments than men on average?

```{r}
dbGetQuery(db,
           "SELECT congress.gender, AVG(posts.comments_count) AS avg_comments
           FROM congress JOIN posts
              ON congress.screen_name = posts.screen_name
           GROUP BY congress.gender")
```

5. For each type of reaction (haha, love, angry...) compute the average count by party. Are there any systematic differences?

```{r}
dbGetQuery(db,
           "SELECT congress.party, AVG(posts.haha_count), AVG(posts.love_count), 
                   AVG(posts.wow_count), AVG(posts.sad_count), AVG(posts.angry_count)
           FROM posts JOIN congress
              ON congress.screen_name = posts.screen_name
           GROUP BY congress.party")
```

6. Compute the total number of "sad" reactions for each domain shared by Republicans. Display the top 10 domains that receive the most "sad" reactions. Then repeat the same for Democrats. What do you find?

```{r}
dbGetQuery(db, 
           "SELECT congress.party, posts.domain, COUNT(*) AS domain_count,
              SUM(posts.sad_count) AS total_sad_reactions
           FROM posts JOIN congress
              ON congress.screen_name = posts.screen_name
           WHERE congress.party = 'Republican'
           GROUP BY posts.domain
           ORDER BY total_sad_reactions
           LIMIT 10")

dbGetQuery(db, 
           "SELECT congress.party, posts.domain, COUNT(*) AS domain_count,
              SUM(posts.sad_count) AS total_sad_reactions
           FROM posts JOIN congress
              ON congress.screen_name = posts.screen_name
           WHERE congress.party = 'Democrat'
           GROUP BY posts.domain
           ORDER BY total_sad_reactions
           LIMIT 10")




dbGetQuery(db,
           "SELECT * FROM posts") # to take a look at posts table
dbGetQuery(db,
           "SELECT * FROM congress")
```






