# Querying a SQL database

Practice writing SQL queries that answer the following questions using the database we created in lecture:

1. Who wrote the post that was shared the most times during this period? What does it say?

```{r}
library(DBI)
db <- dbConnect(RSQLite::SQLite(), "facebook-data/facebook-db.sqlite")

# save tables so that I can have a look at them
dbListTables(db)
congress_tab <- dbGetQuery(db, "SELECT * FROM congress")
posts_tab <- dbGetQuery(db, "SELECT * FROM posts")

dbGetQuery(db, 
           "SELECT screen_name, from_name, message, MAX(shares_count)
           FROM posts")

dbGetQuery(db,
           "SELECT screen_name, from_name, message, shares_count
           FROM posts
           ORDER BY shares_count DESC
           LIMIT 1")

```

2. What about the post with the most comments?

```{r}
dbGetQuery(db,
           "SELECT screen_name, from_name, message, shares_count
           FROM posts
           ORDER BY comments_count DESC
           LIMIT 1")
```

3. Which was the post that received the most likes in January of 2017?

```{r}
dbGetQuery(db,
           "SELECT screen_name, message, likes_count, id
           FROM posts
           WHERE date LIKE '2017-01%'
           ORDER BY likes_count DESC
           LIMIT 1")

```

4. What was the photo posted by a Member of Congress that received the most "angry" reactions?

```{r}
dbGetQuery(db, 
           "SELECT posts.screen_name, message, posts.from_name, posts.type, posts.id, congress.type, angry_count
           FROM posts JOIN congress
              ON congress.screen_name = posts.screen_name
           WHERE posts.type = 'photo'
           ORDER BY angry_count DESC
           LIMIT 1")

```

5. What was the post by `senatorsanders` that received the most "angry" reactions? And the most "love" reactions?

```{r}
dbGetQuery(db,
           "SELECT screen_name, message, type, angry_count, id
           FROM posts
           WHERE screen_name = 'senatorsanders'
           ORDER BY angry_count DESC
           LIMIT 1")

dbGetQuery(db,
           "SELECT screen_name, message, type, love_count, id
           FROM posts
           WHERE screen_name = 'senatorsanders'
           ORDER BY love_count DESC
           LIMIT 1")
```

6. Which was the post mentioning "Trump" that was shared the most? And which was the post that received the most "love" reactions?

```{r}
dbGetQuery(db,
           "SELECT screen_name, message, type, shares_count, id
           FROM posts
           WHERE message LIKE '%Trump%'
           ORDER BY shares_count DESC
           LIMIT 1")

dbGetQuery(db, 
           "SELECT screen_name, message, love_count, id
           FROM posts
           WHERE message LIKE '%trump%'
           ORDER BY love_count DESC
           LIMIT 1")
```





